Toward Centrally Programming Distributed Control Planes

### 2016-01-24 paper/motivation.tex
adding propane example 2 to paper.
Example 2 in paper. Better rule compression for memory reduction.
edits
draft merging of sec 3 and 4

While templates avoids some pitfalls of the fully manual approach, they too are far from ideal. The fundamental issue is the semantic mismatch between desired policies and the level of abstraction of templates. While many policies are network-wide (e.g., prefer customer networks, or never announce a route to a certain prefix to external neighbors), templates are device-level. Operators must still manually decompose network-wide policies into device-level policies that can produce the desired network-wide behavior.
This decomposition is not always straightforward and ensuring policy-compliance can be hard, especially in the face of failures. We illustrate this point using two examples based on policies that we have seen in practice.

@@

@@

We can now see that devising correct configuration for real, larger networks will quickly become a nightmare. These networks will have many neighbors across multiple classes of commercial relationships, differing numbers of links per neighbor, along with several region-, neighbor- or prefix-based exceptions to the default policy. Templates only help to certain extent, e.g., by keeping preference and community values consistent across routers. The operators must still do much of the conceptually difficult work manually, in coming up with the templates themselves. \sysname enables a direct expression of the policies of such networks and compiles them automatically to individual router configurations.

@@

@@

Thus, we see that devising configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname helps operators by implementing their high-level policy specification in a way that guarantees compliance under all failures if that is possible. Otherwise, it generates a compile-time error that informs them that the specification cannot be met. For aggregation, it will also provide a lower bound to operators on the number of failures under which aggregation will not result in blackholes.

>>>>

While templates avoids some pitfalls of the fully manual approach, they too are far from ideal. The fundamental issue is the semantic mismatch between desired policies and the level of abstraction of templates. While many policies are network-wide (e.g., prefer customer networks, or never announce a route to a certain prefix to external neighbors), templates are device-level. Operators must still manually decompose network-wide policies into device-level policies that can produce the desired network-wide behavior.
This decomposition is not always straightforward and ensuring policy-compliance can be hard, especially in the face of failures. We illustrate this point using two examples based on policies that we have seen in practice and we demonstrate the ease with which the
examples are described using our new language \sysname.

@@

\subsection{Example 1:  The Backbone}

@@

We can now see that devising correct configuration for real, larger networks will quickly become a nightmare. These networks will have many neighbors across multiple classes of commercial relationships, differing numbers of links per neighbor, along with several region-, neighbor- or prefix-based exceptions to the default policy. Templates only help to certain extent, e.g., by keeping preference and community values consistent across routers. The operators must still do much of the conceptually difficult work manually, in coming up with the templates themselves.

@@

\para{The Backbone in \sysname}
\sysname simplifies network configuration by allowing users to
specify end-to-end forwarding paths and associating such paths with
the appropriate prefixes.  The \sysname compiler handles the task of 
generating low-level BGP configurations consistent with the user's
end-to-end specifications.  It automatically synthesizes 
import/export filters, local preference, MED attributes, and community tags 
to ensure policy compliance under all possible failure scenarios.

In general, \sysname specifications are written modularly via a series
of user declarations.  
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes}, 
%\emph{paths} and \emph{policies}.  
For example, to begin specification of the backbone
network, we first express the idea that for routes leading to our customer,
we prefer using $R_1$ over $R_2$ (Policy P3 from Figure~\ref{fig:example1}):
\begin{lstlisting}[mathescape=true]
define ExitCust = $exit(R_1 \gg R_2)$
\end{lstlisting}
The above statement defines a set of \emph{ranked paths}.
Included in the set are all paths (and only those paths) that \cd{exit} 
our network
through either router $R_1$ or $R_2$.  The paths that exit through $R_1$
are preferred ($\gg$) to the paths that exit through $R_2$.

To associate ranked paths with
one or more prefixes, we must define a \sysname \emph{policy}.  
Within a policy, statements with the form $\path{x}{p}$
associate the prefixes defined by the predicate $x$ with the set of 
ranked paths defined by the path expression $p$.
For instance, in the following code, ranked paths are associated with
the customer prefixes ($C_{pfx}$), the internal prefixes ($I_{pfx}$),
and all other prefixes (\cd{true}).\footnote{Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements.  Hence, when the prefix predicate \cd{true} follows
statements involving $C_{pfx}$ and $I_{pfx}$, it is interpreted as
$\cd{true} \AND \NOT C_{pfx} \AND \NOT I_{pfx}$.}

\begin{lstlisting}[mathescape=true]
define Routing = {
    $\path{C_{pfx}}{ExitCust \gg end(Cust)}$
    $\path{I_{pfx}}{end(in)}$
    $\path{true}{ExitCust \gg exit(Peer) \gg exit(Prov)}$
}
\end{lstlisting}

Line 2 of the policy above
defines the paths that exit our network to customer prefixes ---
\cd{ExitCust} defines paths through $R_1$ and $R_2$ and in the event
that connections to the customer through both of those routers fail,
a backup route ($end(Cust)$) is defined that admits traffic to go through
any path that ends at the customer.  
Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.  The special keyword \cd{in} represents any location
inside the user's network whereas the keyword \cd{out} represents any location
outside the user's network.
Line 4 applies to any other traffic, and allows for any routes that leave through a peer with a preference for customers over peers over providers. To summarize our progress so far, the \cd{Routing} policy 
implements (P1), (P3) and (P4) from Figure~\ref{fig:example1}

Since, routing still allows transit traffic (e.g., traffic can enter from \textit{Peer} and leave through \textit{Prov}), we can define a new policy that
implements this restriction separately.

\begin{lstlisting}[mathescape=true]
define PP = Peer or Prov
define PPTransit = $enter(PP) \AND exit(PP)$
define CustTransit = $later(Cust) \AND later(Prov)$

define NoTransit = {
    $\path{true}{\neg PPTransit \AND \NOT CustTransit}$
}
\end{lstlisting}

The \textsf{NoTransit} constraint above ensures that requirements (P2) and (P5) are satisfied. In particular, it says that for any prefix, traffic should never both enter and exit the network from \textit{Peer} or \textit{Prov}. Similarly it prevents traffic from ever following paths that leave our network and later passing through both \textit{Prov} and \textit{Cust}.  To implement both \cd{Routing}
and \cd{NoTransit} simultaneously, we simply conjoin them:

\begin{lstlisting}[mathescape=true]
Routing $\AND$  NoTransit
\end{lstlisting}

\sysname will generate per-device import and export filters, local preferences, 
MED attributes, and community tags to ensure that the policy is 
implemented correctly under all failure scenarios.

\subsection{Example 2:  The Data Center}

@@

Thus, we see that devising a configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname helps operators by implementing their high-level policy specification in a way that guarantees compliance under all failures if that is possible. Otherwise, it generates a compile-time error that informs them that the specification cannot be met. For aggregation, it will also provide a lower bound to operators on the number of failures under which aggregation will not result in blackholes.

\para{The Data Center in \sysname}
In our data center example, 
there are primarily three main concerns: 
(1) traffic for the prefix block allocated to each top-of-rack router must be able to reach that router, 
(2) local services must not leak outside the datacenter, and 
(3) aggregation must be performed on global prefixes to reduce churn 
in the network.

\sysname allows us to decompose and specify each of these constraints in a modular fashion. The first constraint is about prefix ownership -- namely, that we only want traffic for certain prefixes to end up at a particular locations. The following definition captures this intent.

\begin{lstlisting}[mathescape=true]
define Ownership = {
    $\path{p_{g1}}{end(A)}$
    $\path{p_{g2}}{end(B)}$
    $\path{p_{l1}}{end(E)}$
    $\path{p_{l2}}{end(F)}$
}
\end{lstlisting}

In English: traffic for prefix $p_{g1}$ is only allowed to follow paths that 
end at router A; traffic not matching $p_{g1}$, but which matches $p_{g2}$ must 
end at router B; and so on.

To capture the second constraint, we can define another task for the core 
routing policy:

\begin{lstlisting}[mathescape=true]
define Routing = {
    $\path{p_{g*}}{any}$
    $\path{p_{l*}}{\NOT enter(out)}$
    $\path{true}{exit(out)}$
}
\end{lstlisting}

The first line states that there is no restriction (\cd{any})
on how traffic must
traverse the network for global prefixes, aside from the default restriction
that traffic must not pass through the user's network and then loop
back on itself. This means traffic for
$p_{g*}$ may be sent either from other routers in the datacenter, or
from external ASes. The second line ensures that traffic for local
prefixes never enters the network from an outside location. This constraint
guarantees that the services remain reachable only to locations
internal to the data center.

As in the backbone example, we can combine these constraints 
logically to specify the network-wide policy.
However, in addition to constraints on the shape of paths, 
\sysname allows the operator to specify constraints on the BGP control plane. 
For instance, a constraint on aggregation is included to ensure that 
aggregation for $p_{agg}$ is performed from locations inside the network 
to locations outside the network. \todo{Say more about aggregation here.
What prefixes are aggregated in to what blocks?  How is this defined?}

\begin{lstlisting}[mathescape=true]
Ownership $\wedge$ Routing $\wedge$ agg($p_{agg}$, $in \rightarrow out$)
\end{lstlisting}

As before, once \sysname compiles the policy, it is guaranteed to hold under 
all possible failure scenarios. In addition, the system can check for 
aggregate-induced black holes in the presence of up to $k$ failures. 
\todo{Note:  Another comment about failure checking for black holes, which may
not be implemented yet.}
